Comparing \pkg{binsegRcpp} with other implementations of binary segmentation for change-point detection

### 2025-05-08 article.Rnw
distributions

We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$, in which we want to find abrupt changes.
We assume that each data point $i\in\{1,\dots,N\}$ has a corresponding weight $w_i>0$.
Binary
segmentation can be interpreted as attempting to find model parameters
$\theta$ which minimize a given weighted loss function $w_i
\ell(\theta, x_i)$. For example, minimizing the square loss $w_i
\ell(\theta, x_i) = w_i (\theta-x_i)^2$ corresponds to maximizing the
normal log-likelihood, when the variance is uniform.

@@

>>>>

We assume there is a sequence of $N$ data $x_1,\dots,x_N\in\mathbb R$,
in which we want to find abrupt changes.  We
further assume that on each segment there is a constant model
parameter $\theta$, typically the mean (and perhaps
variance/scale). We would like to find a sequence of change-points,
such that each segment has a parameter $\theta$ which minimizes a loss
function $\ell(\theta,x)$. The loss functions that we consider to
minimize can be interpreted as maximizing the log likelihood of
several common distributions (Normal, Poisson, Laplace).

Furthermore, we assume that each data point $i\in\{1,\dots,N\}$ has a
corresponding weight $w_i>0$, which is used in the loss function via
$w_i \ell(\theta, x_i)$. For example, minimizing the square loss $w_i
\ell(\theta, x_i) = w_i (\theta-x_i)^2$ corresponds to maximizing the
normal log-likelihood, when the variance is uniform.

@@

\subsection{Distributions and loss functions}

We consider five change-point models with loss functions $\ell$, each
of which have a corresponding value for the \code{distribution.str}
argument of the proposed function, \code{binsegRcpp::binseg}.

\paragraph{Square loss, normal change in mean.}
The total square loss is used for \verb|distribution.str="mean_norm"|.
The segment-specific parameter $\theta\in\mathbb R$ is the mean.

\paragraph{Normal change in mean and variance.}
The total Gaussian negative log likelihood is used as the loss function for \verb|distribution.str="meanvar_norm"|.
The segment-specific parameter $\theta\in\mathbb R^2$ is the mean and variance.
The variance is estimated via
\begin{equation}
  Y/W -\mu(\mu-2X/W).
\end{equation}
TODO discuss max zero var, which is the largest value that should be
considered numerically zero for a variance estimate. \code{INFINITY}
is returned for the loss if the estimated variance is smaller than max
zero var.

\paragraph{Poisson change in rate parameter (mean and variance).}
The total Poisson loss is used for \verb|distribution.str="poisson"|

\paragraph{L1 loss, Laplace change in median.}
The total absolute error is used for \verb|distribution.str="l1"|

\paragraph{Laplace change in median and scale.}
The total Laplace negative log likelihood is used for
\verb|distribution.str="laplace"|
